Teaching Low-Level Concurrency with \pebbles and \landslide

### 2016-08-27 sigcse-17/410.tex
sigcse: save some frames^Wlines and add a conclusion paragraph
sigcse: tom's revisions

\fourten, Operating System Design and Implementation,
is an upper-level undergraduate class
designed for juniors and seniors in Computer Science
% garth removed this, i guess
%and Electrical
and Computer Engineering.
For CS students it is one of five courses which fulfill
a ``systems programming'' degree requirement.
%
The prerequisite course is \twothirteen,
Introduction to Computer Systems,
% TODO CAMREADY
%a course which has been adopted at other
%universities in the U.S.\ and worldwide
adapted from Carnegie Mellon University's 15-213
\cite{sigcse01:CSaPP}.
This section discusses the course learning objectives,
outlines the programming projects,
and mentions some current limitations in the scope
of the course material.

@@

Students who complete the kernel project on time
then work on a kernel-extension project.
Past projects have included
a sound card driver,
a filesystem,
suspend-to-disk,
profiling,
and an in-kernel debugger.
Two more aggressive, projects have been
paravirtualization
%so that their kernels can host guest kernels and
and multi-processor support.
%to their single-processor kernels.

@@

>>>>

\fourten, Operating System Design and Implementation,
is an upper-level undergraduate class
designed for juniors and seniors in Computer Science
% garth removed this, i guess
%and Electrical
and Computer Engineering.
For CS students it is one of five courses which fulfill
a ``systems programming'' degree requirement.
%
The prerequisite is \twothirteen,
Introduction to Computer Systems,
% TODO CAMREADY
%a course which has been adopted at other
%universities in the U.S.\ and worldwide
a course adapted from Carnegie Mellon University %'s 15-213
\cite{sigcse01:CSaPP}.
This section discusses the course learning objectives,
outlines the programming projects,
and mentions some current limitations in the scope
of the course material.

@@

Students who complete the kernel on time
then work on a kernel-extension project.
Past projects have included
a sound card driver,
a filesystem,
suspend-to-disk,
profiling,
and an in-kernel debugger.
Two more aggressive, projects have been
paravirtualization
%so that their kernels can host guest kernels and
and multi-processor support.
%to their single-processor kernels.

@@

\textbf{Time.}
Being a high-level systems elective, \fourten is exceptionally time-consuming,
demanding 20 or more hours weekly during the hardest projects. %thread library and kernel projects.
A more accessible version of the course could be implemented
by extending the time available for the first three projects
and concluding with a stripped-down subset of the kernel,
or by offering the kernel project in an optional second semester.

### 2016-08-27 sigcse-17/intro.tex
sigcse: tom's revisions

Incorporating symmetric multiprocessing (SMP) into the \fourten curriculum would expand the challenges students face in a single semester.
However, we wish to avoid compromising the complexity
%, hardware fidelity,
and learning experience of the project set.
One solution would be to ease the burden of debugging concurrency errors.
Hence, we have built \landslide%
% TODO camready
%\cite{landslide}
, a tool that uses stateless model checking \cite{verisoft,dpor,chess}
as an alternative to the conventional stress tests.
%to force kernel threads to take execution paths that expose concurrency bugs.
\landslide tests student thread libraries fully automatically, and can test kernels provided some manual instrumentation.
It also features an experimental, fully automatic, Pintos testing mode.
We present our experience sharing \landslide with volunteering \fourten students who used it to find bugs in their own code,
and discuss the possibility of integrating it as a main component of the operating systems curriculum.

>>>>

Incorporating symmetric multiprocessing (SMP) into the \fourten curriculum would expand the challenges students face in a single semester.
However, we wish to avoid compromising the complexity
%, hardware fidelity,
and learning experience of the project set.
One solution would be to ease the burden of debugging concurrency errors.
Hence, we have built \landslide%
% TODO camready
%\cite{landslide}
, a tool that uses stateless model checking \cite{verisoft,dpor,chess}
as an alternative to the conventional stress tests.
%to force kernel threads to take execution paths that expose concurrency bugs.
\landslide tests student thread libraries fully automatically, and can test kernels provided some manual instrumentation.
It also features an experimental, fully automatic, Pintos testing mode.
%We present our experience sharing \landslide with volunteering \fourten students who used it to find bugs in their own code,
%and discuss the possibility of integrating it as a main component of the operating systems curriculum.

We describe the \fourten curriculum in detail (\sect{\ref{sec:410}}),
highlight \landslide's design and functionality (\sect{\ref{sec:landslide}}),
present our experience sharing \landslide with volunteering students who used it to find concurrency bugs (\sect{\ref{sec:eval}}),
%and discuss the possibility of integrating it as a main component of the operating systems curriculum (\sect{\ref{sec:warpzone}}).
and discuss ongoing work to incorporate it into the curriculum at \cmu and other schools (\sect{\ref{sec:warpzone}}).
